Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-21 ICML2019/main.tex
add algorithms
update plots
fix bib file

\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS \equiv 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
%
  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $GS$ be the global sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot u(x,r)}{2 GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot GS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{??} proves privacy $\epsilon$ differential privacy of $\expmech$.

@@

Figure \ref{??} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

  \begin{algorithm}
  \caption{$\lexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $LS$ be the local sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot u(x,r)}{2 LS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot LS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot u(x,r)}{2\cdot LS(\dataobs)})}
{\displaystyle\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)\equiv\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix.

@@

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\{0,1\}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
The interested reader will find the proof in Appendix.
We are now ready to define the mechanism $\hexpmech$. $\hexpmech(\dataobs)$ outputs a candidate $r \in \candidateset$ with probability
\[
  \frac {\exp\big(\frac{-\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)}\big)}
  {\displaystyle\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{4\cdot S(\dataobs)})},
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  $\hellinger(\bysinfer(\dataobs),\cdot)$. {\color{red}{why do we choose 1???}}
  \begin{algorithm}
  \caption{$\hexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $S(\dataobs)$ be the vector of $\gamma-$sensitivity of scoring function. 
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon \cdot \ux{r}}{2 S(\dataobs)})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon \cdot}{2\cdot S(\dataobs)})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

@@

\subsection{Privacy of Exponential Mechanism Family}
\subsubsection{$\expmech$ $\epsilon-$Differential Privacy}
\begin{lem}
$\expmech$ is $\epsilon-$differentially private by \cite{dwork2014algorithmic}.
\end{lem}

@@

\subsubsection{Accuracy Bound for Laplace Mechanism}
\label{subsec_accuracy_lap}
\begin{lem}
\label{lem_acc_lap}
Given $Y \thicksim \lap{0}{b}$, the accuracy bound developed for $\lapmech$ is:
\begin{align*}
&
\lapmechPr{\dataobs}{
\begin{aligned}
\lefteqn{\hellinger(\bysinfer(\dataobs), z )}\\ 
&= \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t)
\end{aligned}
}
\\
&\qquad = 
\begin{cases}
\frac{1}{2} (e^{- \frac{\epsilon (t)}{\Delta \bysinfer}} - e^{- \frac{\epsilon (t + 1)}{\Delta \bysinfer}}) &  t \geq 0\\
\frac{1}{2} (e^{\frac{\epsilon (t + 1)}{\Delta \bysinfer}} - e^{\frac{\epsilon (t)}{\Delta \bysinfer}}) & t < 0
\end{cases}
\end{align*}
where $\betad(\alpha, \beta)$ is the true posterior distribution, i.e., $\bysinfer(\dataobs) = \betad(\alpha, \beta)$, and $r_L$ be the posterior produced by Laplace mechanism, i.e., $z = \betad(\alpha + \lfloor Y \rfloor, \beta - \lfloor Y \rfloor)$.
\end{lem}
(Proof is in Appendix III)

@@

\newpage
\section*{Appendix}

>>>>

\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS \equiv 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
%
  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $GS$ be the global sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 GS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(\dataobs, r')}{2\cdot GS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{lem_exp_private} proves privacy $\epsilon$ differential privacy of $\expmech$.

@@

Figure \ref{fig_theory} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

  \begin{algorithm}
  \caption{$\lexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $LS$ be the local sensitivity of scoring function.  
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon\cdot \ux{r}}{2 LS})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot \ux{r}}{2\cdot LS(\dataobs)})}
{\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)\equiv\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix.

@@

\begin{thm}
  \label{thm_gamma_smooth}
  For any $\dataobs, \dataobs'\in\{0,1\}^n$, and $\gamma \in \mathbb{R}^{+}$
  if  $\adj{\dataobs}{\dataobs'}$ then $\frac{1}{S(\dataobs)} - \frac{1}{S(\dataobs')} \leq \gamma$.
\end{thm}
The interested reader will find the proof in Appendix.
We are now ready to define the mechanism $\hexpmech$. $\hexpmech(\dataobs)$ outputs a candidate $r \in \candidateset$ with probability
\[
  \frac {\exp\big(\frac{-\epsilon\cdot \ux{r}}{4\cdot S(\dataobs)}\big)}
  {\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot \ux{r'}}{4\cdot S(\dataobs)})},
\]
where $S(\dataobs)$ is the 1-smooth sensitivity of  $\hellinger(\bysinfer(\dataobs),\cdot)$. {\color{red}{why do we choose 1???}}
  \begin{algorithm}
  \caption{$\hexpmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$. $S(\dataobs)$ be the vector of $\gamma-$sensitivity of scoring function. 
  \STATE \quad {\bf Initialize} a vector $Pr = (0, \dots, 0 )$ of $|\candidateset|$ length.
  \STATE \quad {\bf For} $r$ in $\candidateset$:
  \STATE \quad \quad calculate the outputting probability of $r$.
  \STATE \quad \quad $Pr[r] = 
  {\exp(\frac{\epsilon \cdot \ux{r}}{4 S(\dataobs)})}
/ {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon \cdot \ux{r'}}{4 \cdot S(\dataobs)})}$
\STATE \quad sample $z$ from $\candidateset$ by probability distribution $Pr$.
  \STATE {\bf return} $z$
  \end{algorithmic}
  \end{algorithm}

@@

\subsection{Privacy of Exponential Mechanism Family}
\subsubsection{$\expmech$ $\epsilon-$Differential Privacy}
\begin{lem}
\label{lem_exp_private}
$\expmech$ is $\epsilon-$differentially private by \cite{dwork2014algorithmic}.
\end{lem}

@@

\subsubsection{Accuracy Bound for Laplace Mechanism}
\label{subsec_accuracy_lap}
\begin{lem}
\label{lem_acc_lap}
Given $Y \thicksim \lap{0}{b}$, the accuracy bound developed for $\lapmech$ is:
\begin{align*}
&
\lapmechPr{\dataobs}{
\begin{aligned}
\lefteqn{\hellinger(\bysinfer(\dataobs), z )}\\ 
&= \hellinger(\betad(\alpha, \beta), \betad(\alpha + t, \beta - t)
\end{aligned}
}
\\
&\qquad = 
\begin{cases}
\frac{1}{2} (e^{- \frac{\epsilon (t)}{\Delta \bysinfer}} - e^{- \frac{\epsilon (t + 1)}{\Delta \bysinfer}}) &  t \geq 0\\
\frac{1}{2} (e^{\frac{\epsilon (t + 1)}{\Delta \bysinfer}} - e^{\frac{\epsilon (t)}{\Delta \bysinfer}}) & t < 0
\end{cases}
\end{align*}
where $\betad(\alpha, \beta)$ is the true posterior distribution, i.e., $\bysinfer(\dataobs) = \betad(\alpha, \beta)$, and $r_L$ be the posterior produced by Laplace mechanism, i.e., $z = \betad(\alpha + \lfloor Y \rfloor, \beta - \lfloor Y \rfloor)$.
\end{lem}
(Proof is in Appendix)

@@

\appendix
